```{r}
# Load the necessary packages
library(tidycensus)
library(dplyr)
library(tidyverse)
```

```{r}
census_api_key("7f9fe0f583b0c04f100196af6c16c58796014a1f", install = TRUE, overwrite = TRUE)
readRenviron("~/.Renviron")
```


```{r}

# Load variables for a specific year and geography
# Let's say we want total population (B01003_001) and population under 18 (B01001_003) for all counties in a state

household_children <- get_acs(geography = "state",
                              variables = c("B09005_001E","B09005_002E", "B09005_003E", "B09005_004E", "B09005_005E"),
                              year = 2022,
                              state = "VT",
                              survey = "acs5",
                              output = "wide"
)
```

```{r}
#single householder with minor children(male)
household_children$ratio_male <- household_children$B09005_004E/household_children$B09005_001E
household_children$per_male <- household_children$ratio_male*100 # this should be ratio_male not ratio_m

#single householder with minor children (female)
household_children$ratio_female <- household_children$B09005_005E/household_children$B09005_001E
household_children$per_female <- household_children$ratio_female*100

#two householders with minor children 
household_children$ratio_two <- (household_children$B09005_002E+household_children$B09005_003E)/household_children$B09005_001E
household_children$per_two <- household_children$ratio_two*100

#group median percent
```


```{r}
plotDf <- household_children[,13:ncol(household_children)] |> # we only need column 13 and after
  pivot_longer(1:6,names_to = 'var', values_to = 'value') |>
  # remember that when you plot geom_col, you need the three values to be three observations of the same one variable (that means in the same one column), not three different variables (or columns); then, you need another column as the x which tells us what each value stands for (permale, perfemale, pertwo). Same when you plot geom_bar, except you don't need the column for x.
  filter(var %in% c('per_male', 'per_female', 'per_two')) 
```


```{r}
library(ggplot2)
ggplot(plotDf, aes(x = var, y = value)) + 
  geom_col() +
  labs(x = 'family type', title = "Family types under poverty level")
```

# Change column color

for each column to have a different color (fill), you'll have to tell ggplot what to change colors by (in our case, by var):

```{r}
ggplot(plotDf, aes(x = var, y = value, fill = var)) + 
  geom_col() +
  labs(x = 'family type', title = "Family types under poverty level")
```

You can specify what the colors are. One way is to use a preset.

See more presets here: https://r-graph-gallery.com/38-rcolorbrewers-palettes.html

```{r}
ggplot(plotDf, aes(x = var, y = value, fill = var)) + 
  geom_col() +
  labs(x = 'family type', title = "Family types under poverty level") +
  scale_fill_brewer(palette = 'Spectral')
```

Or you can manually add colors. Like this:

```{r}
ggplot(plotDf, aes(x = var, y = value, fill = var)) + 
  geom_col() +
  labs(x = 'family type', title = "Family types under poverty level") +
  scale_fill_manual(values = c('#9e2a2b','#e09f3e','#0a9396'))
```

## Change background color

there are two ways to do this. The easier one is to pick a pre-set theme from ggthemes. This will change the font and a lot of other things to the pre-set, too.

See all themes here: https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/

```{r}
#install.packages('ggthemes')
library(ggthemes)
ggplot(plotDf, aes(x = var, y = value, fill = var)) + 
  geom_col() +
  labs(x = 'family type', title = "Family types under poverty level") +
  scale_fill_manual(values = c('#9e2a2b','#e09f3e','#0a9396')) +
  theme_tufte()
```

Or you can set the background of the panel to be 'white'. I'm setting it to 'green' to show you what changed after we added this:
```{r}
ggplot(plotDf, aes(x = var, y = value, fill = var)) + 
  geom_col() +
  labs(x = 'family type', title = "Family types under poverty level") +
  scale_fill_manual(values = c('#9e2a2b','#e09f3e','#0a9396')) +
  theme(panel.background = element_rect(fill = 'green'))
```


# How to have plots show up on the right in the "Plots" panel:

This is a settings issue, not a code issue. See: https://posit.co/blog/r-markdown-tips-tricks-1-rstudio-ide/#:~:text=Show%20plots%20in%20the%20Viewer,Voil%C3%A0!

By default, code chunks display R Markdown plots “inline”, or directly underneath the code chunk. If you would rather see the plot in the Viewer pane, go to RStudio > Preferences > R Markdown and unselect “Show output inline for all R Markdown documents”.

Alternatively, just copy all the code (everything in between ``` and ```{r}) into a r script and run it. The plot will show up on the right.